---
title: "Lab 1 - Data visualization"
format: pdf
author: Ayden Frost
editor: visual
---

# Questions

## Part 1

```{r}
library(tidyverse)
```

### Question 1

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100) +
  labs(
    x = "Population Density of Counties",
    y = "Number of Counties", 
    title = "Population Density of Midwesten Counties",
    subtitle = "Binwidth = 100"
  ) 
```

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 1000) +
  labs(
    x = "Population Density of Counties",
    y = "Number of Counties", 
    title = "Population Density of Midwesten Counties",
    subtitle = "Binwidth = 1000"
  ) 
```

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 10000) +
  labs(
    x = "Population Density of Counties",
    y = "Number of Counties", 
    title = "Population Density of Midwesten Counties",
    subtitle = "Binwidth = 10000"
  ) 

```

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100000) +
  labs(
    x = "Population Density of Counties",
    y = "Number of Counties", 
    title = "Population Density of Midwesten Counties",
    subtitle = "Binwidth = 100000"
  ) 
```

The 1000 Binwidth histogram would be ideal as it visualizes the data in a presentable way that is easy to interpret when compared to the other 3 histograms.

### Question 2

```{r}
ggplot(midwest, aes(y = popdensity)) +
  geom_boxplot() +
  labs(
      y = "Population Density of Counties", 
    title = "Population Density of Midwesten Counties",
    subtitle = "Binwidth = 100"
  )

```

The distribution of population sizes amongst counties displayed in both the histogram and boxplot show that most counties tend to be under 12,500 individuals per unit area, with only a small handful of counties breaking the threshold. One county that stands out is Cook county, as it has a population density of roughly 88,000. The most likely reason behind this could be the existence of a city or larger community within the county.

### Question 3

### Question 4

### Question 5

### Question 6

### Question 7

## Part 2

**Enough about the Midwest!**

### Question 8

### Question 9
